---
title: "Data Science e investigação reproduzível e colaborativa"
author: "A Peralta Santos"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
  html_notebook:
    toc: yes
    toc_depth: 2
    toc_float: yes
---

# Hospitals Arrears
This work was developed to explore the arrears of Portuguese Public Hospitals between 2014-2019, all the data in publicly available at ["Portal da Transparencia"](https://transparencia.sns.gov.pt/explore/dataset/divida-total-vencida-e-pagamentos/table/?sort=periodo)

### Set the configurations

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  eval = TRUE
  )

# install.packages("tidyverse")
library(tidyverse) # All purpose library 
```


## Load the dataset
We loaded the databases from the API of the transparency Portal of the Ministry of Health, each time we run the code routine the latest data is loaded. 
```{r}
sns_epe <- readRDS("~/Documents/GitHub/DataScience_wokshop_PT/Data/sns_epe.rds")

sns_sum_m <- readRDS("~/Documents/GitHub/DataScience_wokshop_PT/Data/sns_sum_m.rds")
```


## Create a table with hospitals arrears

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
table_monthly_var<- sns_sum_m %>%
  select(year, month_name, arrears_m, arrear_m_var, percentile_rank) %>%
  filter(year==2019)
```

```{r echo=FALSE,  message=FALSE, warning=FALSE}
library(kableExtra)
kable(table_monthly_var, 
      caption = "Montly Arrears variation of Portuguese E.P.E. Hospitals ", 
      booktabs = T, 
      digits = 1,
      col.names = c("Year",
                    "Month",
                    "Total Arrears",
                    "Monthly Arrears Variation",
                    "Var. Percentile"))  %>%
kable_styling(latex_options = "striped") %>%
  footnote(general = "Agregated monthly arrears of the Portuguese E.P.E. Hospitals")
```




## Vizualize the hospitals arrears

### Arrears over time
```{r}
p1 <- ggplot(data = sns_epe,
            mapping = aes(x = date,
                          y = arrears_sum,
                          colour = year)) +
            geom_point(alpha = 6/10) +
  geom_smooth(method = lm, 
              se = FALSE) +
  geom_vline(xintercept=as.numeric(as.Date("2015-11-01")), colour="grey") + 
  geom_text(aes(x=as.Date("2016-01-01"), 
                label="XXI Governo",y=1000), 
            colour="grey", 
            angle=90, 
            vjust = 1.2,
            size = 3) 


p1 +  labs(title = "E.P.E. Hospitals Arrears",
           subtitle = "2013 - 2018",
           color='Year (linear)',
           y = "Arrears in Million Euros",
           x = "Year",
           caption = "Source: Portal da Transparencia SNS")  + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5)) +
theme_minimal() +
theme(plot.title=element_text( face="bold", 
                               hjust=0.5,
                               lineheight=1.2),
      plot.subtitle=element_text(hjust=0.5)
      )
```

### Arrears Variation
```{r }

sns_sum_m$colour <- ifelse(sns_sum_m$arrear_m_var <= 0, "Negative","Positive")
sns_sum_m$hjust <- ifelse(sns_sum_m$arrear_m_var > 0, 1.3, -0.3)

p1 <- ggplot(data = sns_sum_m,
            mapping = aes(x = date,
                          y = arrear_m_var,
                          label="",
                          hjust=hjust,
                          na.rm = TRUE)) +
            geom_bar(stat="identity",
                     position="identity",
                     aes(fill = colour))  + 
#  geom_text(aes(y=0,colour=colour)) +
  scale_fill_manual(values=c(Positive="#00BFC4",
                             Negative="#F8766D"))



p1 + labs(title = "Monthly variation of Portuguese E.P.E. Hospitals Arrears",
           subtitle = "2014 - 2019",
           y = "Arrears Monthly variation in Million Euros",
           x = "Year",
           caption = "Source: Portal da Transparencia SNS",
           fill = "Arrears Var")  + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
theme_minimal() +
theme(plot.title=element_text( face="bold", 
                               hjust=0.5,
                               lineheight=1.2),
      plot.subtitle=element_text(hjust=0.5)
      ) 
```



```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
sns_sum_m$bailout <- ifelse(sns_sum_m$arrear_m_var < -50,1 ,0 )
sns_sum_m$bailout1 <- ifelse(sns_sum_m$arrear_m_var < -10,1 ,0 )

sns_sum_m<- sns_sum_m %>% 
  arrange(date)  %>% 
  mutate(time = row_number())

sns_sum_m<-sns_sum_m  %>% 
  mutate(bailout_lag1= lag(bailout,1)
         )

sns_bailout <- sns_sum_m %>%
  select(date, bailout, bailout1, bailout_lag1)

sns_model<-full_join(sns_epe, sns_bailout, by = NULL, copy = FALSE)

sns_model<- sns_model %>% 
  arrange(date)  %>% 
  group_by(entidade) %>% 
  mutate(time = row_number())


```

### Arrears by Hospital
```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
p_tile <- ggplot(sns_epe, 
            aes(date, hospital)) +
  geom_tile(aes(fill = m_var_perct_rank),
     colour = "white") + 
  scale_fill_gradient(
    low = "white",
    high = "steelblue") +
  facet_grid(~ year, switch = "x", scales = "free_x", space = "free_x")   +
    #remove x and y axis labels
  labs(y="")+
  #remove extra space
  scale_y_discrete(expand=c(0,0))+
  #define new breaks on x-axis
  scale_x_discrete(expand=c(0,0),
                   breaks=c("2014","2015","2016","2017","2018","2019"))+
    #set a base size for all fonts
  theme_grey(base_size=8)+
  guides(fill=guide_legend(title="Percentile Arrears Var"))+
  labs(x="",y="",
       title="Monthly Arrears Variation of E.P.E. Hospitals",
       caption = "Source: Portal da Transparencia SNS")
  #theme options
  theme(
    #bold font for legend text
    legend.text=element_text(face="bold"),
    #set thickness of axis ticks
    axis.ticks=element_line(size=0.4),
    #remove plot background
    plot.background=element_blank(),
    #remove plot border
    panel.border=element_blank(),
    plot.title=element_text( face="bold", 
                               hjust=0.5,
                               lineheight=1.2)
  )

```

```{r}
print(p_tile)
```

# Gross Domestic Product and Life Expactancy 
_From the Kieran Healy's Book "Data Visualization"_


In this example we will use the data from Gapminder website and explore the association between GDP and Life Expectancy

```{r include=FALSE}
library(gapminder)
gapminder <- gapminder

model_colors <- RColorBrewer::brewer.pal(3, "Set1")
model_colors
```

```{r echo=FALSE , fig.align="center", fig.cap="Figure 3: GDP and Life Expectancy "}

p0 <- ggplot(data = gapminder,
            mapping = aes(x = log(gdpPercap), 
                          y = lifeExp))

p1 <- p0 + geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", 
                aes(color = "OLS", 
                    fill = "OLS")) +
    geom_smooth(method = "lm", 
                formula = y ~ splines::bs(x, df = 3),
                aes(color = "Cubic Spline", 
                    fill = "Cubic Spline")) 


p1 + scale_color_manual(name = "Models", values = model_colors) +
    scale_fill_manual(name = "Models", values = model_colors) +
    theme(legend.position = "right") + labs(title = "Life expectancy and GDP per capita",
           y = "Life expectancy in years",
           x = "Log of GDP per capita")
```

```{r include=FALSE}
out <- lm(formula = lifeExp ~ gdpPercap + pop + continent,
          data = gapminder)

out1 <- lm(formula = lifeExp ~ gdpPercap,
          data = gapminder)
```


```{r include=FALSE}
library(broom)
library(Hmisc)
library(devtools)
library(socviz)

out_comp <- tidy(out)

out_conf <- tidy(out, conf.int = TRUE)


out_conf <- subset(out_conf, term %nin% "(Intercept)")
out_conf$nicelabs <- prefix_strip(out_conf$term, "continent")

```

We can create a table for the regression summary results 


```{r }

table1 <- out_conf %>% 
  select(term, estimate, conf.low, conf.high, p.value)  %>% 
  filter((str_detect(term, "continent")))

table1 <- table1 %>%
 dplyr::mutate(term = recode(term, 
     "continentAmericas" = "America",
     "continentAsia" = "Asia",
     "continentEurope" = "Europe",
     "continentOceania" = "Ocenia")
  )

kable(table1, 
      align = "c",
      digits = 2,
      col.names = c("Continent",
                           "Estimate",
                           "Low 95% CI",
                           "High 95% CI",
                    "P value"),
      caption = "Table. Linear regression of the effect of GDP on Life expectancy") %>%
  kable_styling(full_width = F)
```



### THE END


